Combining checkpointing and data compression\\ for high frequency seismic imaging

### 2018-10-03 paper/paper.tex

Our performance model assumes
 that the compression algorithm behaves uniformly
across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
now becomes
\begin{equation}
\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} +
t_c) + \mathbf{N}_R \cdot (\frac{\mathbf{S}}{\mathbf{F}\mathbf{B}} + t_d),
\end{equation}
where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
and $t_d$ are compression and decompression times, respectively. At the same
time, the recomputation overhead decreases
because $\mathbf{F}$ times more checkpoints are now available.

TODO add figure here that shows memory on x-axis and speedup on y axis. Split into 3 regions. Region 1, 
where Revolve+Compression makes sense, Region 2, where compression would probably not bring in enough benefit, 
and region 3 where compression would enable the computation to be done without revolve. 
\section{Results}

\section{Conclusions and Future work}

@@

\section*{Acknowledgment}

TODO (Acknowledgments for Navjot)

This work was supported by the U.S. Department of Energy, Office of Science,
Office of Advanced Scientific Computing Research, Applied Mathematics and
Computer Science programs under contract number DE-AC02-06CH11357.
We would also like to acknowledge the support from the SINBAD II project and
the member organizations of the SINBAD Consortium.

>>>>

Our performance model assumes
 that the compression algorithm behaves uniformly
across the different time steps of the simulation, i.e. that we get the same compression ratio, compression time and 
decompression time, no matter which of the $N$ possible checkpoints we try to compress/decompress. The storage overhead
now becomes
\begin{equation}
\mathbf{O}_{SR} = \mathbf{N}_W \cdot (\frac{\mathbf{S}}{\mathbf{F}
  \cdot \mathbf{B}} + t_c) + \mathbf{N}_R \cdot
(\frac{\mathbf{S}}{\mathbf{F} \cdot \mathbf{B}} + t_d)
\end{equation}
where $\mathbf{F}$ is the compression ratio (i.e. the ratio between the uncompressed and compressed checkpoint), and $t_c$
and $t_d$ are compression and decompression times, respectively. At the same
time, the recomputation overhead decreases
because $\mathbf{F}$ times more checkpoints are now available.

\section{Results}
Figure \ref{fig:varying_memory} shows the speedups predicted by the
performance model for varying amount of available memory. A big
shortcoming that was realised in ZFP was the lack of parallelization
in the decompression functions. The ZFP manual talks about the
difficulty in implementing a parallel mode decompressor. To estimate
the benefit from parallel decompression, we scaled the serial
decompression time by the same factor that we saw between the serial
and parallel compressors on a given platform. We called this scaled
decompression time, ``theoretical decompression time'' and show this
on the plot with a dotted line. The
right-most region of this plot has been the focus of many studies
mentioned earlier in this paper. The extension to the regions with
lower memory is new here and we see that the maximum benefit is gained
in the most memory-constrained scenarios (region 1), while there is a
region where Revolve-based recomputation alone would be more
beneficial (region 2). As expected, compression alone (with no
Revolve) dominates region 3. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-memory.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  memory. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The dotted curve represents the hypothetical speedup
if ZFP supported parallel decompression. The two vertical lines at 20GB and 52GB clearly
divide the plot into 3 regions. The left-most region, with less than
20GB memory sees the maximum benefit from compression. The middle
region, with memory between 20GB and 52GB, is better off using only
Revolve and no compression. The right-most region, with memory greater
than 52GB represents the scenario where the entire history of the
wavefield can fit in memory post-compression. In this case compression
eliminates the use of Revolve - hence the speedup. This is the region
that has been well-studied in past work.}
\label{fig:varying_memory}
\end{figure}

Figure \ref{fig:varying_compute} shows the speedups predicted by the
performance model for kernels of different computational cost. Perhaps
unsurprisingly, the benefit gained by compression drops rapidly if the
time to compute a single timestep is less than the time to compress
the computed data (since recomputing would be cheaper in that case). For
increasing computational cost, the speedup from compression saturates
at a bound. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-compute.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  compute cost. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. The benefits of compression drop rapidly if the
computational cost of the kernel that generated the data is much lower
than the cost of compressing the data. For increasing computational
costs, the benefits are bounded.}
\label{fig:varying_compute}
\end{figure}

Figure \ref{fig:varying_nt} shows the speedups predicted for differing
number of timesteps. It is interesting to note that there is a
consistent increase in the speedup as the number of timesteps goes
up. 
\begin{figure}
\begin{center}
\includegraphics[width=\linewidth]{images/varying-nt.pdf}
\end{center}
\caption{The speedups predicted by the performance model for varying
  number of timesteps to be reversed. The baseline
(1.0) is the performance of a Revolve-only implementation under the
same conditions. It can be seen that compression becomes more
beneficial as the number of timesteps is increased.}
\label{fig:varying_nt}
\end{figure}
\section{Conclusions and Future work}

@@

\section*{Acknowledgment}
This work was funded by the Intel Parallel Computing Centre at
Imperial College London and EPSRC EP/R029423/1 . 
This work was supported by the U.S. Department of Energy, Office of Science,
Office of Advanced Scientific Computing Research, Applied Mathematics and
Computer Science programs under contract number DE-AC02-06CH11357.
We would also like to acknowledge the support from the SINBAD II project and
the member organizations of the SINBAD Consortium.
